---
title: "Joins and k-nn"
author: "Keaton Wilson"
date: "2/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Context  
We're running short on tracking collars for leopard seals after pulling all of 
the collars from Budget Collars (they seem to be failing more often, and their 
battery life is terrible - putting our fish-harvesting team in danger).  

One of our intrepid data science team members found an old box of collars and 
some data on all of the collars (it was a real challenge getting this data 
off of an old floppy drive, but she managed). So we're going to spend the next
lessons tackling two tasks: first, we're to work to join data sets together. 
The main goal here is to create a single data set that includes the data 
we have previously looked at for collars, the data on the new collars, as well 
as the additional data on the old collars.  

This is tricky, as we we cant just do simple rbinds and cbinds like we've done
in the paste - the collar IDs need to be matched up with their counterparts 
across datasets, and new collars need be added, but with the column that 
contains maker ID missing.   

Second, we'll dive into some machine learning. We'll talk a bit about what 
machine learning is (and isn't) and then use a frequently used classification 
algorithm, k-nearnest neighbors to predict which maker made which of our 
unidentified collars.  

## Joins  
Being able to join together data from related tables is a key skill in data 
science, and for working with larger data structures (databases with their 
own languages, like SQL).  

## Group Discussion - Joins in R
Take 5 minutes to read through this [article](https://medium.com/@HollyEmblem/joining-data-with-dplyr-in-r-874698eb8898) 
individually. https://tinyurl.com/rfmnvcs

After this 5 minutes, discuss the questions below with your group. We'll then 
discuss them as a class. Make sure that someone in the group (a reporter) can 
report out answers to each of the questions below.  

1. What's the difference between a left join and an inner join?  
2. Can you think of a reason you would use an anti-join?  
3. What type of join should we use to bind the three tables we're working on 
with our seal collar data?  
4. Can you draw a daigram that illustrates what a left-join does?  


# Working through our data  
Let's load in the data we're working with and work through implementing 
our joins.  

```{r}
library(tidyverse)

#loading in the data
collars = read_csv("https://tinyurl.com/sp7b25x")
new_collars = read_csv("https://tinyurl.com/t3of8rk")
old_collars_new_data = read_csv("https://tinyurl.com/ufawycq")





`

