---
title: "Module 3-1"
author: "Keaton Wilson"
date: "2/6/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction to the Problem and Summarizing Data  

## The challenge  
Part of the food that folks are being fed with are fish that are caught, not 
farmed (as in the previous module). These fish are caught by members of team 
antarctica, but there is extreme risk involved (think Deadliest Catch, but 
colder).  

One of the main hazards to these teams are leopard seals - they're apex 
predators in this ecosystem, and can cause serious injury, especially when large 
schools of fish are involved (a researcher was killed by a Leopard Seal in 
Antarctica in 2003).  

One way we've worked to remove some of this danger is by working with large 
mammal researchers to place radio collars on a number of seals that live in 
areas that are typically fished. Fishing boats are equipped with radios that can
detect the presence of seals in the area and avoid high-risk sites.  

In recent months, a problem has arisen. Some of the collars are failing, 
leading to some very close calls - one team member was pulled into the water 
after attempting to untangle a net when a seal lunged at some fish trapped in 
the net.  
 
The large mammal team (they work on orcas as well as leopard seals) replaces 
collars frequently, based on the manufacters reccomendations for battery life 
and general wear and tear, but it seems as if some of the collars are dying 
earlier than expected and putting our team in danger. 

We've been tasked with determining why the units are failing and if we can tie 
it to a particular manufacturer (we've sourced radio collars from two different
companies).  

## The Data  

Our data on collars can be accessed here: https://tinyurl.com/sp7b25x  

Let's explore and talk about the data a bit:  
1. Two manufacturers  
2. Battery life - The average number of days a particular collar lasts (this is
recorded in the unit and stored when the battery dies)  
3. Signal distance - the maximum signal distance that a particular collar was 
recorded at.  
4. Fail - collars that have failed in the past (e.g. they've been recovered 
from seals that were noticed by the team but that didn't ping the radio 
equipment).  

## Group Challenge 1 - Summarizing
We covered how to do grouped summaries in R in Module 1 (and practiced a bit
in module 2), while also covering how to write your own custom functions in 
module 2. It's time to put this together. Write your own custom function that 
does the following:  
1. Summarizes the data by calculating the mean, min and max for battery life 
and signal distance, while counting up the number of failures. You'll want to
do this for both manufacturers. 
```{r}
library(tidyverse)

collars = read_csv("https://tinyurl.com/sp7b25x ")
summary_1 = collars %>%
  mutate(fail = as.logical(fail), 
         maker = as.factor(maker)) %>%
  select(-collar_id) %>%
  group_by(maker) %>%
  summarize_if(is.numeric, list(min = min, max = max, mean = mean)) %>%
  select(-contains("collar_id"))

summary_2 = collars %>%
  mutate(fail = as.logical(fail), 
         maker = as.factor(maker)) %>%
  select(-collar_id) %>%
  group_by(maker) %>%
  summarize_if(is.logical, sum)

final = bind_cols(summary_1, summary_2) %>%
  select(-maker1)

```

## Visualizing the Data - an introduction to a better way to visualize the data
So far, we've used the base R plotting syntax, but there is another way. It's 
part of the tidyverse and its from a package called ggplot2.  

Instead of explaining the syntax to you, we're going to do an in-class reading 
assignment.  

[Article to read](https://cfss.uchicago.edu/notes/grammar-of-graphics/)  

Spend 10 minutes reading the article above (take notes!). When the time is up, 
I'll signal you to get into your groups and discuss the article. There is a lot 
of material here - make sure that you cover what the grammar of graphics is, 
what the different components are, and how they relate to code structure and 
syntax. We'll be iteratively building some ggplot code on the data above, and 
I'll be asking groups to explain each part as we go - so be prepared!  

```{r}
ggplot(collars, aes(y = signal_distance, x = battery_life, col = maker)) +
  geom_point()
```

