---
title: "module_3_2"
author: "Keaton Wilson"
date: "2/12/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Plotting and Means Comparisons  
Last class, we examined the collars data set, read a bit about the grammar of 
graphics and analyzed some code for a ggplot of some of the data we've been 
working with.  

Let's take a stab at making our ggplot from scratch.  

In your groups, work together to develop code that utilizes the collars data
[https://tinyurl.com/sp7b25x](https://tinyurl.com/sp7b25x) and plots:  
1. Two histograms one for battery life and one for signal length  
2. Color coded for each manufacturer  
3. Correct and appropriate labels  
4. A vertical line that plots the median of each group (this is trickier - 
remember to use the google)  

```{r}
library(tidyverse)
collars = read_csv("https://tinyurl.com/sp7b25x")

collars %>%
  ggplot(aes(x=battery_life, fill = maker)) +
  geom_histogram(alpha = 0.6, bins = 20) +
  geom_vline(data = collars %>%
               group_by(maker) %>%
               summarize(median_battery = median(battery_life)), 
             aes(xintercept = median_battery, col = maker)) +
  ylab("Count") +
  xlab("Battery Life") +
  scale_color_discrete(name = "Maker") +
  scale_fill_discrete(name = "Maker") +
  theme_classic()

collars %>%
  ggplot(aes(x=signal_distance, fill = maker)) +
  geom_histogram(alpha = 0.6, bins = 20) +
  geom_vline(data = collars %>%
               group_by(maker) %>%
               summarize(median_signal = median(signal_distance)), 
             aes(xintercept = median_signal, col = maker)) +
  ylab("Count") +
  xlab("Signal Distance") +
  scale_color_discrete(name = "Maker") +
  scale_fill_discrete(name = "Maker") +
  theme_classic()


```
**What is another way to plot these data?**  

## Stepping back and thinking about the problem  
Ok, so stepping back - what is the fundamental question we're asking here?  

Pose this to the class.  

**Fundamentally, we want to know if there a difference between the two 
manufacturers in these two variables - battery life and signal distance**  

They look different in the graphs you made, but is this enough? What do I mean?  

Draw three different histograms on the board.  
![histogram comparison](../assets/histogram_example.pdf)

How do we tell whether there is an **actual** or **meaningful** difference in 
the means?  

## Introduction to a t-test  
There is a formalized set of statistical tools to answer this question. 
Some of you may have heard of t-tests before - which is what we use on this 
type of data.  

Let's briefly talk through the logic here.  
1. Our data are a sample of a larger population (think about all of the collars
ever produced for both companies).  
2. What we're really interested in is the the difference in the means between 
the two groups. If they're the same, this difference is 0. If they're different, 
the difference is something larger or smaller than 0.  
3. the code is simple, but the output is a bit trickier:  

```{r}
t.test(battery_life ~ maker, data = collars)

```
4. Breaking down the information here:  
  a. *t*: is the test statistic - here, it's a metric of how different the 
  difference of the two means is from 0. High (or low) values indicate a big 
  difference, small values are closer to 0.  
  b. *df*: is degrees of freedom - it's a measure of your sample size and some 
  other stuff - not something we're really going to focus on here.  
  c. *p-value*: a measure of certaintanty of the difference outlined in the 
  *t-statistic* above, it is veryyyyyy small. This means that there is an 
  extremely low probability that obtaining a difference in means of 0 is very,
  very unlikely. There are a lot of benchmarks in different fields for what this
  value should be below to consider the difference *significant*, typically 
  < 0.05 is considered significant.  
  d. *95 percent condience interval*: this is the range that we can expect the 
  test statistic (difference in means) to fall in 95% of the time given the data.  

## Exploring a t-test on your own
Work on your own (but feel free to consult your group) to run a t-test on the 
other variable we are interested in (signal distance). Be able to describe each 
piece of the output, what it means, and the overall finding for the test. 
Also be able to relate it back to the figures we generated above. 


