---
title: "Joins and k-nn"
author: "Keaton Wilson"
date: "2/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Context  
We're running short on tracking collars for leopard seals after pulling all of 
the collars from Budget Collars (they seem to be failing more often, and their 
battery life is terrible - putting our fish-harvesting team in danger).  

One of our intrepid data science team members found an old box of collars and 
some data on all of the collars (it was a real challenge getting this data 
off of an old floppy drive, but she managed). So we're going to spend the next
lessons tackling two tasks: first, we're to work to join data sets together. 
The main goal here is to create a single data set that includes the data 
we have previously looked at for collars, the data on the new collars, as well 
as the additional data on the old collars.  

This is tricky, as we we cant just do simple rbinds and cbinds like we've done
in the paste - the collar IDs need to be matched up with their counterparts 
across datasets, and new collars need be added, but with the column that 
contains maker ID missing.   

Second, we'll dive into some machine learning. We'll talk a bit about what 
machine learning is (and isn't) and then use a frequently used classification 
algorithm, k-nearnest neighbors to predict which maker made which of our 
unidentified collars.  

## Joins  
Being able to join together data from related tables is a key skill in data 
science, and for working with larger data structures (databases with their 
own languages, like SQL).  

## Group Discussion - Joins in R
Take 5 minutes to read through this [article](https://medium.com/@HollyEmblem/joining-data-with-dplyr-in-r-874698eb8898) 
individually. https://tinyurl.com/rfmnvcs

After this 5 minutes, discuss the questions below with your group. We'll then 
discuss them as a class. Make sure that someone in the group (a reporter) can 
report out answers to each of the questions below.  

1. What's the difference between a left join and an inner join?  
2. Can you think of a reason you would use an anti-join?  
3. What type of join should we use to bind the three tables we're working on 
with our seal collar data?  
4. Can you draw a daigram that illustrates what a left-join does?  


# Working through our data  
Let's load in the data we're working with and work through implementing 
our joins.  

```{r}
library(tidyverse)

#loading in the data
collars = read_csv("https://tinyurl.com/sp7b25x")
new_collars = read_csv("https://tinyurl.com/t3of8rk")
old_collars_new_data = read_csv("https://tinyurl.com/ufawycq")


#explore 
glimpse(collars)
glimpse(new_collars)
glimpse(old_collars_new_data)

# Tell me about these data, how are we going to join them?
```

### Step 1. The join

```{r}
old_collars_full_data = collars %>% left_join(old_collars_new_data, 
                                              by = c("collar_id", "maker"))

```


### Step 2. a bind or a join? 
So now we need to put the new collars in here...

How do we join this? Is it even a join at all?

```{r}
full = bind_rows(old_collars_full_data, new_collars)
```


## Prediction, Machine Learning  
Now, we're going to work through an example of machine learning, which is 
used in a variety of ways in data science - and across industries and 
discplines.  

The goal here is that we have ~ 50 unlabeled collars, and we want to be able to 
predict who made them. Remember that Budget Collars LLC provided us with collars 
that are failing, which is a risk to human life! So we need to throw out any 
collars that we think were made by this company, and only incorporate new 
collars from our unlabeled batch that we think were made by Collarium Inc. 

### What is machine learning? Group reading and discussion.  
Take 10 minutes to read [this article](https://medium.com/@randylaosat/a-beginners-guide-to-machine-learning-dfadc19f6caf). 

I'll signal when 10 minutes is up. Get into your groups and discuss the following
questions.  
1. What is machine learning? Is it different than AI?  
2. What types of general problems do machine learning techniques tackle?  
3. Given what you've read, what kind algorithm type might we employ for our 
problem?  

## knn in R
K-nearest neighbors is a classification algorithm (e.g. it attempts to predict 
labels that are classes in R). We're going to work through a small example on 
our data set before attempting to build the full model. 

### Check-in  
What are our classes/labels here?  

### Data cleaning and prepping, splitting and normalizing

First, we need to process our data a bit before we jump into our algorithm.  

```{r}
# remind ourselves of what full contains
glimpse(full)


# First, we want to filter out the missing data - we'll run our full algorithm 
# on this later. 

full_filtered = full %>%
  filter(!is.na(maker))

collars_to_label = full %>%
  filter(is.na(maker))


# Next, let's get rid of some unneeded columns
filtered = full_filtered %>%
  select(-collar_id, -fail,-antenna_length, -weight) %>% 
  mutate(maker = as.factor(maker))
```

### Check-in  
Why are we doing this? We're keeping it to two variables here... later on we'll
do more!  

On to normalization.  
```{r}

# custom function for normalizing
nor <-function(x){
  (x -min(x))/(max(x)-min(x))
}

knn_data = filtered %>%
  mutate(battery_life_norm = nor(filtered$battery_life),
         signal_distance_norm = nor(filtered$signal_distance)) %>%
  select(-battery_life, -signal_distance)

```

### Check-in  
Why are we doing this? Can you think of a good reason why we have to normalize 
the values here? What does normalize mean?  

### We have data that we're ready to put into our algorithm - but first we need 
to figure out how the algorithm actually works. 

We're going to watch a brief video: https://www.youtube.com/watch?v=MDniRwXizWo

After we're finished, get into your groups and discuss the following points:  
1. For our small, filtered data set with two variables, how would you draw a 
simple diagram like in the video that demonstrates how k-nn is working?  
2. Can you envision how this would work/what it would look like with more than 
two variables?  
3. What would happen if you increase or decrease K in the GoT example above, 
or generally?  

## Let's try our hand at k-nn

```{r}
#loading package to do knn
#install.packages('class')
library(class)

#splitting the data up 
train = as.data.frame(knn_data[,-1])
cl = knn_data[,1] %>% pull()

test = expand.grid(x=seq(min(train[,1]-0.5), max(train[,1]+0.5),
                            by=0.1),
                      y=seq(min(train[,2]-0.5), max(train[,2]+0.5), 
                            by=0.1))

mod_test = knn(train, test = test, cl, k = 7, prob=TRUE)


```



