---
title: "Module 4_1"
author: "Keaton Wilson"
date: "4/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A new road - bypass the penguins please  
The overarching goal of this study builds on the last. We're looking at 
constructing a new road that gives the fishing team access to some new sites 
that seem to be relatively leopard-seal free. However, the road crosses through
some Gentoo penguin nesting grounds. We have data on a lot of different sites, 
and want to pick a route that minimizes the impact to the penguins (e.g. 
we want to pick sites that currently have low nesting success).  

We're going to examine our first data set of the module: 
https://tinyurl.com/w9tol83  

Let's explore this data set before moving on to the focus of this lesson.  
Load in the data and use the tools we've learned to explore the data. Answer 
the following questions:  
1. What does each column represent  
2. What questions do you have about the data?  

```{r}
library(tidyverse)

penguins = read_csv("https://tinyurl.com/w9tol83")

head(penguins)
glimpse(penguins)

```

## Bootstrapping  
What would be nice is to have the average number of grass tussocks for each 
site. We can do this with the group_by and summarize tools we've been working 
all semester.  

```{r}
penguins %>%
  group_by(site) %>%
  summarize(avg_tussocks = mean(avg_tussocks), 
            n = n())

```

One big question though - how confident are we that these means represent the 
actual mean for a given site? E.g. we didn't sample every square meter of all 
of these sites. We took a **sample**, and we assume that the mean here is a 
good approximation of the actual mean. We can use confidence intervals to give 
us a little more information about the spread of this value and how confident 
we are. 

**Check in and ask about where we have used confidence intervals before**  

## By the bootstraps  

Bootstrapping is inherently about resampling without replacement. We take a 
small data set (all data sets are small, right?) that we assume is 
representative of a population of samples, and draw out samples with 
replacement, calculate some statistic (a mean, here), and do it over again. 
Many, many times. 

We can use these samples to create a confidence interval of the statistic 
(the mean), or even a comparison of means like a t-test.  

### Video explanation  
let's take ~10 minutes to watch the following video: [https://www.youtube.com/watch?v=-YgeLJRZQYY](https://www.youtube.com/watch?v=-YgeLJRZQYY).  
After we've finished, get into your groups and discuss the following 
questions (and designate a reported to report out to the class):  
1. What is the utility of bootstrapping many many times (> 10k)?  
2. Why do we sample with replacement?  
3. How might we apply this to our data if we wanted to construct means (and confidence intervals) for each site?  

## Building a loop to do bootstrapping for us  
Let's build some code to run a bootstrap for the overall mean number of 
tussocks across all sites 

```{r}
# length vector - how many samples do we have?
n = length(penguins$avg_tussocks)
# number of bootstraps
B = 1000
# Making an empty vector to pipe values into
result = rep(NA, B)

# looping
for (i in 1:B){
  boot.sample = sample(n, replace = TRUE)
  result[i] = mean(penguins$avg_tussocks[boot.sample])
}
# calculating upper and lower bounds
lower = mean(penguins$avg_tussocks) + (-1*2*sd(result))
upper = mean(penguins$avg_tussocks) + (1*2*sd(result))
CI = c(lower, upper)

print(CI)
```


## Group challenge Every site! Expanding the bootstrap  
Take the code above and modify/expand it to get a series of means and confidence intervals for each site. Hint: think about turning it into a function!

```{r}
sites = unique(penguins$site)
summaries = list()
for(i in 1:length(sites)){
  df_tmp = penguins %>% filter(site == sites[[i]])
  # length vector - how many samples do we have?
  n = length(df_tmp$avg_tussocks)
  # number of bootstraps
  B = 1000
  # Making an empty vector to pipe values into
  result = rep(NA, B)
  
  # looping
  for (j in 1:B){
    boot.sample = sample(n, replace = TRUE)
    result[j] = mean(df_tmp$avg_tussocks[boot.sample])
  }
  # calculating upper and lower bounds
  lower = mean(df_tmp$avg_tussocks) + (-1*2*sd(result))
  upper = mean(df_tmp$avg_tussocks) + (1*2*sd(result))
  summary = data.frame(name = sites[i], 
                       mean = mean(df_tmp$avg_tussocks), 
                       lower = lower, 
                       upper = upper)
  summaries[[i]] = summary
}

summary_df = bind_rows(summaries)
```

## Visualizing these data  
Let's finish this mini-project by plotting the means and bounds of our confidence intervals so visualize the set of differences among all the sites.   
```{r}
summary_df %>%
  ggplot(aes(y = mean, x = as.factor(name), group = as.factor(name))) +
  geom_point() +
  geom_point(aes(y = lower), color = "blue")


```




