---
title: "module_4_3"
author: "Keaton Wilson"
date: "4/22/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## More hypothesis testing and Extensions of t-tests and permutation tests (ANOVA)  

Let's start working with a new data set that complements what we've been 
working with so far. 

### New data  
```{r}
library(tidyverse)
site_data = read_csv("https://tinyurl.com/yardhofj")
glimpse(site_data)
site_data

```
A big one is making comparisons between the 1971 data and 
the 2011 data. Can you think of some questions we could ask?

### Hypothesis brainstorm  
Take a few minutes and write down some ideas you have for questions that could 
be asked and associated hypotheses?

## Focusing on one research question  
One big question we might ask is whether or not there have been changes in the 
number of tussocks between the two time periods. Do we have some statistical 
tools in our toolkit to answer this?  

1. Visualization  
2. t-test (think about the difference of this dataset though)
3. Permutation test (think about the difference of this dataset though)

## Group Discussion and Work 
Spend 10 minutes in your groups putting together one (or more) of these tools 
that demonstrates an answer. You should be able to explain the output and 
differences in laymans terms.  

```{r}
library(tidyverse)
site_data %>%
  ggplot(aes(x = tussocks, fill = as.factor(year))) +
  geom_density(alpha = 0.8) +
  xlab("Number of Tussocks per Site") +
  ylab("Density") +
  scale_fill_discrete(name = "Year") +
  theme_classic()


```