---
title: "module_4_2"
author: "Keaton Wilson"
date: "4/16/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Site-specific data, permutation tests  

We're going to use another data set on penguins for the rest of this module. 
Remember that there are a few goals here:  
1. To explore the penguin nesting data set and generate some conclusions about 
what factors are related to nesting success  
2. Take these conclusions to inform our decisions about where the newly 
developed road should go. We're trying to mitigate environmental impact as much 
as possible by understanding what factors drive penguin nesting success.  

## Exploring the new data  
Let's take a look at the new data set. 
```{r}
library(tidyverse)
penguins_2 = read_csv("https://tinyurl.com/yardhofj")

glimpse(penguins_2)
```

## Independent Exploration of Data  
Spend 3-5 minutes on your own exploring the data set. Focus on the following 
questions:  
1. How many individual sites are represented here?  
2. What is similar from the first data set we examined?  
3. What is different from the first data set we examined?  
4. Are there any columns/variables that you have questions about?  

## Comparisons of means - another approach... 

One of the main questions we're interested in with these data is whether or not 
there is a difference in our variables across all sites but between time 
intervals (e.g. have things changed between 1971 and 2011). Today, we're going 
to specifically focus on number of tussocks. What tools have we talked about so 
far that would address this question?  

We're going to explore another option today that is a little more flexible - 
there are a bunch of assumptions associated with t-tests that can really mess 
things up - and the type of test we're working on today gives us a way to 
still make comparisons, even when some of those assumptions aren't met. This 
method also allows us to go over hypothesis testing, p-values and some other
statistical fundamentals that are part of the data science toolkit.  

## The permutation test - but with alpacas  
Check out this link: 
[https://www.jwilber.me/permutationtest/](https://www.jwilber.me/permutationtest/)  

Spend about 5 minutes on your own walking through the logic and example and then
work in your groups to answer the following discussion questions (remember to 
designate someone as the reporter for reporting out!):  
1. What is the conclusion in the example scenario(from a reality-perspective, not a 
statistical one) - what actionable insights were gained from this experiment?  

2. Can you explain what the big difference is between this method and a t-test?  

3. Can you give a brief outline of each step for the permutation test?  

4. Can you think of any problems with running through all the permutations 
possible, as the author talks about?  

5. What is the hypothesis we want to test with our data? What is the null 
hypothesis associated with this question?  

### Coding permutation tests in R  
```{r}
library(perm)
glimpse(penguins_2)

test = perm::permTS(x = penguins_2 %>% 
                  filter(year == 1971) %>%
                  pull(tussocks), 
             y = penguins_2 %>%
                  filter(year == 2011) %>%
                  pull(tussocks))

```

### Practice - does stone size change with year?  
By yourselves, generate code that tests the following hypothesis: there is a 
difference in the mean stone size between 1971 and 2011.  

What is the null hypothesis? What does the output of the permutation test 
tell us?  


