---
title: "Module 4_4"
author: "Keaton Wilson"
date: "4/29/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Bringing it all together - how do multiple factors affect penguin nesting?  

So, we're going to spend the next two classes focusing on what is probably the 
most used tool in statistics, and address the big question we've been dancing 
around for most of this module: **what factors predict penguin nesting, and can 
we build a model that allows to pick out sites that are the least destructive to
the penguin population if we develop a road there**?  

## Visualizing the problem - Two variables  
Let's begin by reading in our data and building a plot that explores some 
of the variables of interest.  
```{r}
library(tidyverse)

# importing the data
site_data = read_csv("https://tinyurl.com/yardhofj")
glimpse(site_data)

# filtering so we just have the most current data set at our dispsoal
site_data_2011 = site_data %>% 
  filter(year == 2011)

# plotting
ggplot(site_data, aes(x = tussocks, y = num_nests)) +
  geom_point() +
  theme_classic()

```
Ultimately, the goal of linear regression is to be able to draw a line through 
this cloud of points that best represents the pattern between the number of 
tussocks and the number of nests. This line represents a simple model of how we 
think the system works and allows us to make predictions about the number of 
nests at a given site given we know how many tussocks are at that site.  

How do we choose this line? Let me illustrate the problem.   
```{r}
ggplot(site_data, aes(x = tussocks, y = num_nests)) +
  geom_point() +
  geom_abline(slope = 0, intercept = 20, color = "blue") +
  geom_abline(slope = 2.4, intercept = 9, color = "green") +
  theme_classic()
```

## Individual Exploration - Least Squares?  
Spend 10 minutes searching the web. Your goal is to be able to explain how 
linear regression works. How do we figure out what the best line is? What is 
this process called, and how would you explain it to someone who has never taken 
a statistics course before?  

Be ready to report out!  

## Visualizing the regression line  
So now that we have a basic idea of how this works, let's explore how we 
can visualize a linear regression and also dive into some nitty gritty about 
building a model and interpreting what R tells us about the model.  

```{r}
# do a regular smooth first
ggplot(site_data, aes(x = tussocks, y = num_nests)) +
  geom_point() +
  geom_smooth() +
  theme_classic()

ggplot(site_data, aes(x = tussocks, y = num_nests)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_classic()

```

# Building a real model and exploring/interpreting  

So the plot above is great, but... it's all happening within ggplot. We can't
extract any useful information about the regression there. For that, we have to
use some built-in code to generate a model (think of it as the equivalent of the 
t.test material we did earlier).  

```{r}
# specifying our model  
tussock_mod = lm(num_nests ~ tussocks, data = site_data)
summary(tussock_mod)

```

## Group discussion and explanation of output  
Spend 10 minutes in your groups discussing the following questions:  
1. Can you explain what the following components are and how to interpret them:  
  a. Estimate for Intercept and tussocks  
  b. Pr(>|t|)  
  c. Adjusted R-squared  
  d. p-value  
2. What is the overall conclusion about this model?  

This is a big lift! Try and divide and conquer within your groups, and remember 
that Google is your friend.  

As always - be ready to report out!  


  