---
title: "Settlers of Antarctica - Module 1 - Day 4"
author: "Keaton Wilson"
date: "12/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Day 4 - 2-dimensional data and introduction to the tidyverse  

## Learning Objectives  
1. Students will practice programming fundamentals like installing and loading packages, using helpfiles and navigating file structure.  
2. Students will be able to use the fundamental functions of the tidyverse: select, filter, mutate, summarize and the pipe.  
3. Students will be able to compare base R methodology and tidyverse methodology for filtering and indexing dataframes.  
4. Students will able to use functions above to gain insight from real-world data.  

## Tidyverse - what is it?  
Differenet programming languages have different syntax... and tidyverse is a package (or set of packages) offered in R that have similar goals and a set of unified syntax designed specifically to work with 2-dimensional data. When I started learning R, none of this was available, but the ecosystem has changed a lot in the last 10 years, and this is the set of tools I use to work with data every day.   

## Getting the tidyverse (and what's in it) - Group Challenge and discussion

Let's begin by scoping out what is in the tidyverse. Go to www.tidyverse.org and browse the site. In your groups, answer the 3 questions below, and be prepared to report out at the end:   
1. What are the core packages in the tidyverse?  
2. Give a one sentence summary about what each package does.  
3. What is one question you have about the tidyverse/core packages?  

## Practice with the tidyverse  

We're going to use a real data set on climate change from Berkeley. It outlines temperatures in major cities across the world since 1750. We're going to use this data set to learn the tidyverse!  

First things first, let's load it up. 

```{r}
library(tidyverse)

# download.file("https://tinyurl.com/vvsykhm")

climate_df = read_csv("../data/GlobalLandTemperaturesByMajorCity.csv")

# call the climate data... this looks a bit different
climate_df

# It's a tibble! Like a dataframe but better. :) Lots of information to dissect here. Let's look at another way to quickly view our data.  

glimpse(climate_df)


# 
```
## Select()ing columns  

Let's use our first function, select.  

```{r}
# Select at its core allows you pick out a selection of columns from your data
# Can folks remember how do this with base R?

climate_df$dt
climate_df[,1:2]

#Select does this, but with more power. 
select(climate_df, dt)

# Multiple columns
select(climate_df, dt, City, Country)
select(climate_df, dt:Country, -City)

# Similar process for rows - though we typically don't have rownames in R
slice(climate_df, 1:4)
```

## Filtering a Dataframe with filter()  

`filter()` allows you filter rows by certain conditions. Recall that we did this a bit with base R.  

```{r}

# base R
climate_df[climate_df$Country == "United States",]

# filter
filter(climate_df, Country == "United States")

# Pretty similar, but easy to chain stuff together
filter(climate_df, Country == "United States" & AverageTemperature > 25)

# Also worth noting here that we haven't saved any of this. We need to write to a new object. 

us_df = filter(climate_df, Country == "United States")

```

## Group Challenge  
Work with the climate data we've using this class period. Construct a small set of code that does the following:  
1. Slims down the full dataframe to one that contains the columns dt, AverageTemperature and City.  
2. Filters the data for Paris and Mexico City (hint:this may be hard than you think...remember to do quality control on your data) with an average temperature less than 22.  
3. Name this new dataframe "cold_paris_and_mexico_city"

```{r}
# Piped
cold_paris_and_mexico_city = climate_df %>%
  select(dt, AverageTemperature, City) %>%
  filter(City == "Paris" | City == "Mexico", 
         AverageTemperature < 22)

# Not piped
slim = select(climate_df, dt, AverageTemperature, City)
filtered = filter(slim, City == "Paris" | City == "Mexico", 
                  AverageTemperature < 22)
cold_paris_and_mexico_city = filtered
```