---
title: "Module 2-1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Learning Objectives  
1. Students will become familiar with the real-world problem for this module - 
food poisoning due to an outbreak of disease in one of the species of fish that 
is grown in recirculating aquaculture systems.  
2. Students will be be able to generate two types of plots using base R syntax 
to visualize single variables (histograms and distributions) and two variables 
(scatterplots).
3. Students will develop visual-thinking skills to create visualizations that 
allow them to explore patterns in data, draw inferences and create solutions.  

## Introduction to the problem  
Overall, the issue here is that we have a wave of people getting sick across 
the team. People are coming in complaining of stomach sickness - doctors have 
ruled out a communicable viral infection like norovirus, and it seems to be a 
food contamination issue.  

The two main sources of food that are grown on site and distributed to team 
members are [plants grown in hydroponic greenhouses](https://thespoon.tech/scientists-in-antarctica-grow-food-without-soil-or-sunlight-get-ready-mars/). 
(mostly swiss chard, cucumbers and radishes) and fish (tilapia, a tolerant warm
-water species).  

**Diagram this next part on the board**  

We have a good set of data on team members that are sick, and how much fish or 
plant material they incorporate into their diets (team members vary in the 
composition of their diet - people are allowed to choose how much of different 
food sources they use).  

## Discussion - how might we figure this out?  
Spend 5 minutes brainstorming in your groups how you might figure out whether 
plants or fish are the culprits? Be ready to report out.  

## Introduction to the dataset and visualization in R  

First we're going to pull in the data and give it a quick inspection/exploration
before we start to work through some of the visualization tools in R.  

```{r}
library(tidyverse)

sick = read_csv("https://tinyurl.com/unavf2v")
sick
glimpse(sick)

# See any problems? Looks like there might be some age data that is messed up 
# some of the age values are wayyyy too young. Let's exclude all of those. 

sick = sick %>%
  filter(age > 18)

# Basic visualization - in base R, the plot(), and some others function is your 
# friend. We'll talk about some tidyverse ways to plot that are a bit more 
# powerful later, but you really can do any kind of plotting you want with base 
# R. 

hist(sick$age)

```

## Mini Group Discussion - Histograms  
What is this showing us? What conclusions can we draw from this data 
visualization? How could this visualization be improved?   

## Group Project - make a better histogram.  
Work in your groups to build a histogram of weights with properly labeled axes, 
a vertical line that depicts the median, and an appropriate visually appealing 
bin size. Check out the `hist()` help page, and the `abline()` function. 

```{r}
hist(sick$age, breaks = 15, xlab = "Age", ylab = "Frequency", main = NULL)
abline(v = median(sick$age))

```

## Brainstorm - Group activity  
We've covered one type of visualization that just shows one variable... but what
we're really interested in is figuring out if fish or plants are the culprit in 
food poisoning. Spend 10 minutes in your groups sketching out a visualization 
that might give us insight into this.  

## Scatter plots  
```{r}
plot(x = sick$age, y = sick$perc_fish)
```

Not super informative, because there isn't really a relationship between a 
person's age and the amount of fish they consume. At least not in this sample. 

# Building and interpreting
In your group, build out the plot that you conceived earlier that determines 
whether there is a correlation between the percentage of fish eaten and the 
number of trips to the doctor in the past 6 months. Use proper labelling and 
aesthetic design principles to make it as visually appealing as possible.  

```{r}
plot(x = sick$perc_fish, y = doctor_trips, 
     xlab = "Percentage Fish in Diet", 
     ylab = "Number of Trips to Doctor", 
     main = NULL, 
     pch = 16)

```
