---
title: "Module 2-2 - Functions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Exploration  
So, it does seem like there is something wrong with the fish, given the 
correlation we found earlier. Let's check out our tank data. 

```{r}
library(tidyverse)

fish_tank_data = read_csv("https://tinyurl.com/tbyskxl")
glimpse(fish_tank_data)
```

## Discussion of the problem  

So here, the scenario, we have this data, and we know what the critical 
temperature cutoffs are for both species, but the problem is that they're in 
farenheit... and the values in our data frame are in celsius. The fish 
biologists we worked with on this were from the UK and US, and for some reason 
didn't give us information the appropriate format. 

Anything below 59ª F for Trout and anything below 65º F for Tilapia is 
critically low, and can result in suppression of the immune system. 

So here is the first challenge, how do we use this information given our current 
data frame. Ultimately, we want a filtered data frame that just shows us tanks 
that are below these values... trying to figure out how common it is.  

With your groups, brainstorm some approaches you think might work. 
Don't actually code them out, just diagram, psuedo code or talk through them. 
As always, be ready to share.  


## Custom Functions  
There are many solutions to this problem, but one that is particularly useful 
and powerful in R is building your own functions. We've already used functions a
ton, and you can write your own... to do absolutely anything you want. This has 
lot of uses... but the big one here is that we can write a simple function that 
converts celsius to farenheit and apply it to our entire data set, in just a few 
lines of code.   

```{r}
