---
title: "Module 2_2 - Functions"
author: "Keaton Wilson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Learning Objectives  
1. Students will be able to recall some fundamental components of the biology 
of recirculating aqauculture systems and udnerstand how these components link 
to the problem at hand - particularly linking environmental tank conditions to 
physiology which in turn can affect disease rates.  
2. Students will practice preliminary data explorations on tank data.  
3. Students will be able to create custom functions that perform simple tasks 
(converting values in celsius to farenheit), with an eye to the utility of this 
generalizeable programming technique.  
4. Students will practice data summary techniques from module 1, particularly 
tidyverse syntax. 

## Data Exploration  - query to the class? Where did we leave off last time?
So, it does seem like there is something wrong with the fish, given the 
correlation we found earlier. Let's check out our tank data. 

```{r}
library(tidyverse)

fish_tank_data = read_csv("https://tinyurl.com/tbyskxl")
glimpse(fish_tank_data)
```

## Discussion of the problem  

So here, the scenario, we have this data, and we know what the critical 
temperature cutoffs are for both species, but the problem is that they're in 
farenheit... and the values in our data frame are in celsius. The fish 
biologists we worked with on this were from the UK and US, and for some reason 
didn't give us information the appropriate format. 

Anything below 59ª F for Trout and anything below 65º F for Tilapia is 
critically low, and can result in suppression of the immune system. 

So here is the first challenge, how do we use this information given our current 
data frame. Ultimately, we want a filtered data frame that just shows us tanks 
that are below these values... trying to figure out how common it is.  

With your groups, brainstorm some approaches you think might work. 
Don't actually code them out, just diagram, psuedo code or talk through them. 
As always, be ready to share.  


## Custom Functions  
There are many solutions to this problem, but one that is particularly useful 
and powerful in R is building your own functions. We've already used functions a
ton, and you can write your own... to do absolutely anything you want. This has 
lot of uses... but the big one here is that we can write a simple function that 
converts celsius to farenheit and apply it to our entire data set, in just a few 
lines of code.   

```{r}
# Here is the basic syntax

new_sum = function(x,y){
  z = x+y
  return(z)
}

new_sum(2,4)
```

## Building a custom function for converting celsius to farenheit  
In your groups, build a custom function called `c_to_f()` that converts a 
single value of celsius to farenheit. Check out the web for the equation on how 
to do this. Be prepared to show your code and run some test values.  

```{r}
c_to_f = function(c = NULL){
  f = (c*1.8)+32
  return(f)
}

c_to_f(c = 40)
```

## Thinking about iteration and looping  

So, we have a great function that we can use to convert our celsius values to 
farenheit, but how do we apply that to everything... we've built the function 
for one value, not for a vector of values...  

Quickly in your groups, brainstorm some ideas on how you might implement this. 

## Three main ways to iterate  
1. for loops  
2. apply functions  
3. Re-write your function  

Next time we'll cover some options, but for now... the solution is actually 
really simple. The function we built actually will auto-iterate over a vector...
because R is amazing at doing vector stuff. Check it out. 

```{r}
c_to_f(fish_tank_data$avg_daily_temp)

# So, we can do something like this
fish_tank_data = fish_tank_data %>%
  mutate(avg_daily_temp_F = c_to_f(avg_daily_temp))

fish_tank_data

```

## Filtering based on our new converted column  
Ok, so we've solved one problem, but not the main one... who remembers the 
original question we set out to examine here?  

**How many tanks are below the temperature cutoffs for each species?**

Take the rest of the class period to work on figuring this out (hint: remember 
the group_by and filter and summary functions we used in module 1?). If you can't finish, 
work in groups remotely to figure out a solution by the beginning of the next 
class.  

```{r}
fish_tank_data %>%
  filter(species == "tilapia" & avg_daily_temp_F < 75 |
         species == "trout" & avg_daily_temp_F < 59) %>%
  group_by(species) %>%
  summarize(n = n())

```
  
