---
title: "Module 2_4 - Simulation and Iteration"
author: "Keaton Wilson"
date: "1/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Review  
Let's take a minute to review as a class what we've talking about, what the 
problem is and where we're at with it.  
1. What is the main problem we're trying to address?  
2. What did we cover last time in class?  
3. Where did we leave off?  

Ultimately, we have a contaminated food supply that we think is making people
sick. We've noticed a correlation between the amount of fish in someone's diet 
and now we're using simulation to compare our expectations for disease in the 
tanks to what we measured in our small (50 tank sample).   

Last time, we worked through building a function to test this, but we realized 
it has limited utility because it only generates one mean... and this mean 
changes every time, because of the nature of random draws from a distribution.  

## Iteration  
Today we're going to dive into one of the most important topics in programming, 
iteration. Something we touched on before, but haven't really explored. The 
ultimate goal here is to build a way to run through our function multiple times 
and calculate a 'mean of means', which should give us a better overall picture 
of what our expected value is.  

## loops  
The most basic way to do iteration is with loops. It's essentially telling R to 
keep doing things a certain number of times, and then doing something with the 
outputs for each iteration.   

```{r}
# Let's use a basic example, but one which we would never actually use, as R
# does basic vectorized stuff like this by default. 
# Let's generate a logarithm of each value in a vector and output this to a new 
# vector. 

vec_in = c(2, 6, 8, 9, 5)

for(i in 1:length(vec)){
  log(vec[i])
}

# Ok... but nothing happened....
# We have to set up some other stuff first

# output vector to feed into
vec_out = c()
for(i in 1:length(vec_in)){
  x = log(vec[i])
  vec_out[i] = x
}
vec_out
```

Things to remember about loops:  
1. Remember to initialize an object before the loop to feed into  
2. Proper syntax - curly braces!  
3. Iterating through pieces with the variable you're using (i, in the example 
above).  

## Applying loops to our simulation function  

We're now going to run our simulation function 100 times. Quick check in - why 
are we doing this?!  

```{r}
# loading in data
library(tidyverse)
fish_tank_data = read_csv("https://tinyurl.com/tbyskxl")

# function
fish_sick_simulator = function(mean = NULL, species = "trout", sd = NULL){
  if(species == 'trout'){
    sick_fish_vec = round(rnorm(n=17, mean = mean, sd = sd))
    return(mean(sick_fish_vec)/74)
  } else {
    sick_fish_vec = round(rnorm(n=33, mean = mean, sd = sd))
    return(mean(sick_fish_vec)/100)
  }
}

# Double check our function to make sure it's still working  
# We're working on Tilapia first! 

fish_sick_simulator(mean = 4, species = "tilapia", sd = 2)

# Let's loop it!  
sim_vec = c()
for(i in 1:100){
  sim_vec[i] = fish_sick_simulator(mean = 4, 
                                   species = "tilapia", 
                                   sd = 2)
}

sim_vec
mean(sim_vec)

```

## Group Challenge  
Construct a similar simulation for trout. Be prepared to answer the following
questions:  
1. Do the sickness percentages measured by the aquaculture team line up with our
expecations? 
2. Are there differences between the two species? What does this point to?  

